A DPG method for compressible flow problems

### 2012-07-24 docs/Jesse/PECOStalk/proposalIntro.tex

Our aim is to develop a \emph{stable, higher order adaptive scheme} for the steady compressible laminar Navier-Stokes equations in transonic/supersonic regimes that is \emph{robust for very small viscosity}.  In particular, we hope to present a method for which $hp$-adaptivity can be applied to problems in compressible flow.  The goal of this dissertation will be to develop a mathematical theory demonstrating the robustness in $\epsilon$ of our method for singularly perturbed convection-diffusion problems, and to demonstrate its feasibility as a CFD solver by applying it to several benchmark problems.  
\todo{Check with LD if it's what he's looking for}

@@

We will introduce the SUPG method at an abstract level for illustrative purposes only.  Further details and perspectives on the SUPG method can be found in \cite{SUPG}, as well as in an upcoming book by Hughes.  The convection-diffusion equation can be written as follows:
\[
Lu = \left(L_{\rm adv} + L_{\rm diff} \right) u = f,
\]
where $L_{\rm adv}u \coloneqq \div \left(\beta u\right)$ is the first order advective operator, and $L_{\rm diff}u \coloneqq \epsilon \Delta u$ is the second-order diffusive operator.  Let us assume $u$ to be a linear combination of piecewise-linear basis functions $\phi_i, i = 0,\ldots,N$ (then, within each element, $L_{\rm diff} u = 0$).  The SUPG method is then to solve
\[
\int_\Omega L u v + \int_\Omega \tau \left(L_{\rm adv}v\right)\left( L u - f\right) = \int_\Omega f v 
\]
for where $\tau$ is the SUPG parameter.  For uniform meshes in 1D, $\tau$ is chosen such that, for $f=0$, the matrix system resulting from SUPG is exactly equal to the finite difference system under ``exact" artificial diffusion.  However, unlike exact artificial diffusion, for $f\neq 0$, the SUPG method still delivers optimal stabilization.  In fact, the SUPG finite element solution in 1D is nothing less than the nodal interpolant and the best $H^1_0$ approximation of the exact solution, as seen in Figure~\ref{fig:SUPG}.

@@

The method can be made to ``work" for $p>1$, but the results are not as strong as for linear elements, and there is no clear generalization of SUPG to higher order elements.  Additionally, the stabilization properties of SUPG are weaker in higher dimensions, and the SUPG solution is no longer the $H^1$ best approximation for 2D and 3D problems \todo{find citation}.  Despite these caveats, SUPG is still the most popular stabilization method of choice for convection-diffusion type problems, in both academic and industry applications.

@@

This addition of residual-based stabilization terms can be interpreted as a modification of the test functions as well.  For SUPG, the formulation can equivalently be written as
\[
\int_\Omega\left( Lu-f\right) v_i = \int_\Omega f v_i, \quad i = 0,\ldots,N
\]
\todo{check to make sure this is correct}
where $v_i$ is defined as
\[
v_i = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
In other words, the test function $v_i$ is a perturbation of the basis function $\phi_i$ by a scaled advective operator applied to $\phi_i$.  For a linear $C^0$ basis function (the ``hat" function), this naturally leads to a bias in the upwind or streamline direction of the flow $\beta$, as seen in Figure~\ref{fig:SUPG}.

>>>>

Our aim is to develop a \emph{stable, higher order adaptive scheme} for the steady compressible laminar Navier-Stokes equations in transonic/supersonic regimes that is \emph{robust for very small viscosity}.  In particular, we hope to present a method for which $hp$-adaptivity can be applied to problems in compressible flow.  The goal of this dissertation will be to develop a mathematical theory demonstrating the robustness in $\epsilon$ of our method for singularly perturbed convection-diffusion problems, and to demonstrate its feasibility as a CFD solver by applying it to several benchmark problems.

@@

We will introduce the SUPG method at the abstract level for illustrative purposes only.  Further details and perspectives on the SUPG method can be found in \cite{SUPG}, as well as in an upcoming book by Hughes.  The convection-diffusion equation can be written as follows:
\[
Lu = \left(L_{\rm adv} + L_{\rm diff} \right) u = f,
\]
where $L_{\rm adv}u \coloneqq \div \left(\beta u\right)$ is the first order advective operator, and $L_{\rm diff}u \coloneqq \epsilon \Delta u$ is the second-order diffusive operator.  Let us assume $u$ to be a linear combination of piecewise-linear basis functions $\phi_i, i = 0,\ldots,N$ (then, within each element, $L_{\rm diff} u = 0$).  If $b(u,v)$ and $l(v)$ are the bilinear form and load for the standard Galerkin method (resulting from multiplying by a test function $v$ and integrating both convective and diffusion terms by parts), the SUPG method is then to solve $b_{\rm SUPG}(u,v) = l_{\rm SUPG}(v)$, where $b_{\rm SUPG}(u,v)$ and $l_{\rm SUPG}(v)$ are defined as
\begin{align*}
b_{\rm SUPG}(u,v) &= b(u,v)+ \sum_{K} \int_{K} \tau \LRp{L_{\rm adv} v} \LRp{L u - f} \\
l_{\rm SUPG}(v) &= l(v) + \sum_{K} \int_{K} \tau \LRp{L_{\rm adv}v} f
\end{align*}
for where $\tau$ is the SUPG parameter. For uniform meshes in 1D, $\tau$ is chosen such that, for $f=0$, the matrix system resulting from SUPG is exactly equal to the finite difference system under ``exact" artificial diffusion.  However, unlike exact artificial diffusion, for $f\neq 0$, the SUPG method still delivers optimal stabilization.  In fact, the SUPG finite element solution in 1D is nothing less than the nodal interpolant and the best $H^1_0$ approximation of the exact solution, as seen in Figure~\ref{fig:SUPG}.

@@

For appropriately chosen $\tau$, the method can be generalized for higher order elements as well.  In higher dimensions, and the SUPG solution is very close to, but no longer the $H^1_0$ best approximation for 2D and 3D problems \cite{HughesSangalliSUPG}.  Since its inception, SUPG is and has been the most popular stabilization method of choice for convection-diffusion type problems, in both academic and industry applications.

@@

This addition of residual-based stabilization terms can be interpreted as a modification of the test functions as well.  For SUPG, the formulation can equivalently be written as
\[
b\LRp{u,\tilde{v}_{i}} = l\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
In other words, the test functions $\tilde{v}_{i}$ is a perturbation of the basis function $\phi_i$ by a scaled advective operator applied to $\phi_i$.  For a linear $C^0$ basis function (the ``hat" function), this naturally leads to a bias in the upwind or streamline direction of the flow $\beta$, as seen in Figure~\ref{fig:SUPG}.
